---
title: "Mayo_CBE_RNASeq_SETUP"
author: "Stella Belonwu"
date: "06/08/2021"
output: html_document
---
Goal: Select APOE 33, 34, and 44 samples and corresponding meta and count data from reprocessed Mayo Cerebellum Lobe RNA-Seq Data\
https://www.synapse.org/#!Synapse:syn8690799 \
https://www.synapse.org/#!Synapse:syn3817650 for covariates\

Load datasets
```{r}
setwd("/home/Box/Bulk_RNASeq_Data/Mayo_CBE")
mayo_covar<- read.csv("MayoRNAseq_RNAseq_CER_covariates.csv",header= TRUE, sep = ",", stringsAsFactors=FALSE)
mayo_data<- read.table("Mayo_CBE_all_counts_matrix.txt.gz",header= TRUE, sep = "\t", stringsAsFactors=FALSE)
ProteinCodingGenes<- read.csv("ensembl_ProteinCodingGenes.csv",header= TRUE, sep = ",", stringsAsFactors=FALSE)
```

Load necessary libraries
```{r}
library(DESeq2) 
library(tidyverse)
library(tximport)
library(ggrepel)
library(magrittr)
library(dplyr)
```

Format metadata: Select AD & Control, Remove ApoE 23,24
```{r}
#Remove rows with missing values
length(unique(mayo_covar$SampleID)) #278
mayo_covar$SampleID[is.na(mayo_covar$Source)] #7 missing 
mayo_covar$SampleID[is.na(mayo_covar$Tissue)] #7 missing 
mayo_covar2<- mayo_covar[!(is.na(mayo_covar$Tissue)),]

#Check RIN/ quality
range(mayo_covar2$RIN) #all above 5, all good

#Select for AD and Controls
length(unique(mayo_covar2$Diagnosis)) #4
unique(mayo_covar2$Diagnosis) #[1] "PSP"              "AD"               "Control"                      "Pathologic Aging"
table(mayo_covar2$Diagnosis)
mayo_covar3<- mayo_covar2[mayo_covar2$Diagnosis == "AD" | mayo_covar2$Diagnosis == "Control",]

#Add "X" to make metaID names match dataID names
mayo_covar3$SampleID<- paste0("X", mayo_covar3$SampleID)

#Remove ApoE 22, 23, 24
table(mayo_covar3$ApoE) #23(14),24(1),33(92),34(47),44(5)
class(mayo_covar3$ApoE)

mayo_covar4<- mayo_covar3[mayo_covar3$ApoE == "33"|mayo_covar3$ApoE == "34"|mayo_covar3$ApoE == "44" ,]
table(mayo_covar4$Diagnosis) #78ad, 66c
table(mayo_covar4$ApoE) #33(92),34(47),44(5)

#Pick columns needed for analysis
colnames(mayo_covar4)
mayo_meta<- mayo_covar4[,c(1,5:9)]  #add 9 for flowcell
colnames(mayo_meta)[6]<- "Batch"

#Fix Age and make in decades 50s, 60s,70s,80s,90s
range(mayo_meta$AgeAtDeath)
unique(mayo_meta$AgeAtDeath)
class(mayo_meta$AgeAtDeath)
mayo_meta2<- mayo_meta

mayo_meta2$Age<- as.numeric(mayo_meta2$AgeAtDeath)
mayo_meta2<- mayo_meta2 %>%
  mutate(Age= ifelse(AgeAtDeath >= 50 & AgeAtDeath <= 59,"50s", Age)) %>%
  mutate(Age= ifelse(AgeAtDeath >= 60 &AgeAtDeath <= 69,"60s", Age)) %>%
  mutate(Age= ifelse(AgeAtDeath >= 70 & AgeAtDeath <= 79,"70s", Age)) %>%
  mutate(Age= ifelse(AgeAtDeath >= 80 & AgeAtDeath <= 89,"80s", Age)) %>%
  mutate(Age= ifelse(AgeAtDeath == "90_or_above","90s", Age)) 
  
unique(mayo_meta2$Age)
mayo_meta2$Age %<>% factor
rownames(mayo_meta2)<- mayo_meta2$SampleID
mayo_meta2$AgeAtDeath<- NULL
mayo_meta2$ID<- NULL
```

Format count data: Select protein coding genes & Select samples present in metadata
```{r}
#Select protein coding genes
mayo_data2<- mayo_data[-c(1:4),]

class(mayo_data2$feature) #character
mayo_data2$feature<-gsub("\\..*","",mayo_data2$feature) #remove numbers/characters after "." to match ensembl gene ids from biomart
length(unique((mayo_data2$feature))) #60275 genes

rownames(mayo_data2)<- mayo_data2$feature
sum(rownames(mayo_data2) %in% ProteinCodingGenes$ensembl_gene_id) #count number of protein coding genes 19679/60275
mayo_data3<- mayo_data2[mayo_data2$feature %in% ProteinCodingGenes$ensembl_gene_id,] #select only protein coding genes
 
#Select samples in final meta table
mayo_data4<- mayo_data3[,c(colnames(mayo_data3) %in% rownames(mayo_meta2))]
```

Match order of meta and count sample names
```{r}
summary(rownames(mayo_meta2) %in% colnames(mayo_data4)) #144 TRUE
summary(rownames(mayo_meta2) == colnames(mayo_data4)) #140 FALSE, 4 TRUE

match(colnames(mayo_data4),rownames(mayo_meta2)) #returns the position of the matches (indices) with respect to the second vector
mayo_idx<- match(colnames(mayo_data4),rownames(mayo_meta2))
mayo_meta3<- mayo_meta2[mayo_idx,] #rearrange rows
summary(rownames(mayo_meta3) == colnames(mayo_data4)) # 144 TRUE

#Save subsetted dataframes  
setwd("/home/Box/Bulk_RNASeq_Data/Mayo_CBE")
write.csv(mayo_data4, file="Mayo_CBE_data.csv")
write.csv(mayo_meta3, file="Mayo_CBE_meta.csv")
```

Exploratory Data Analysis on Mayo Dataset to be analyzed
```{r}
table(mayo_meta3$Diagnosis) # AD 78, Control 66
table(mayo_meta3$Sex) # F 76, M 68
table(mayo_meta3$ApoE) #33(92),34(47), 44(5)
table(mayo_meta3$Batch) #15 in total
table(mayo_meta3$Age) #50s (1), 60s(12), 70s(26), 80s(73), 90s(32)
table(mayo_meta3$ApoE[mayo_meta3$Diagnosis == "AD"]) #33(35), 34(38), 44(5)
table(mayo_meta3$ApoE[mayo_meta3$Diagnosis != "AD"]) ##33 (57), 34 (9)
table(mayo_meta3$Sex[mayo_meta3$Diagnosis == "AD"]) #F 45, M 33
table(mayo_meta3$Sex[mayo_meta3$Diagnosis != "AD"]) #F 31, M 35
table(mayo_meta3$Sex[mayo_meta3$Diagnosis == "AD"],mayo_meta3$ApoE[mayo_meta3$Diagnosis == "AD"]) 
table(mayo_meta3$Sex[mayo_meta3$Diagnosis != "AD"],mayo_meta3$ApoE[mayo_meta3$Diagnosis != "AD"])
```
